Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-09-15 manuscript/manuscript.tex
enhanced discussion

Our study highlights the vital importance of robust data curation in
maintaining the reliability and utility of open datasets, with a close
examination of the NYC 311 data. As these datasets are applied across
diverse fields such as public services and academic research, issues
like inconsistencies, missing values, and formatting errors can
significantly undermine the quality of insights. Trustworthy machine
learning systems, which rely heavily on high-quality data, are
particularly vulnerable to errors when data is not properly
curated. For example, biases or inconsistencies in input data can
compromise both the accuracy and fairness of predictive models
\citep{rahm2000data}.

This research further emphasizes the necessity of harmonizing data
from multiple city agencies, particularly in resolving discrepancies
in field names, formats, and definitions. Trustworthy machine learning
depends on not only accurate and clean data but also consistent data
across sources. This challenge becomes more acute in datasets spanning
long time periods, where changes in agency structures or reporting
standards can introduce subtle biases. As \citet{rahm2000data} noted,
ensuring long-term data consistency is crucial for longitudinal
studies and predictive modeling, as even minor shifts in data can
result in significant deviations in outcomes.

Finally, the intersection of data curation and machine learning for
public policy applications opens new opportunities for enhancing
governance. By improving data quality, machine learning models can
more effectively predict trends, allocate resources, and identify
issues such as service delays. The COVID-19 pandemic demonstrated the
critical role of real-time, trustworthy data in responding to public
needs. Without proper data curation, the accuracy and timeliness of
machine learning insights would have been compromised, affecting
crucial decision-making processes. Future efforts could focus on
automating data curation within real-time data pipelines, ensuring
that the data feeding into machine learning models remains both clean
and reliable \citep{chu2016data}.

>>>>

Our study highlights the critical role of robust data curation in
ensuring the reliability and utility of open datasets, as demonstrated
through the examination of the NYC 311 data. As these datasets are
applied across diverse fields such as public services and academic
research, inconsistencies, missing values, and formatting errors can
significantly undermine the insights derived from them. Trustworthy
machine learning systems, which rely heavily on high-quality data, are
especially vulnerable to such issues. Errors in the underlying data,
such as biases or inconsistencies, can compromise both the accuracy
and fairness of predictive models, impacting real-world decisions in
urban planning and policy \citep{rahm2000data, geiger2020garbage}.

This research further emphasizes the necessity of harmonizing data
across city agencies, particularly in resolving discrepancies in field
names, formats, and definitions. Consistency in data is essential for
trustworthy machine learning, where inconsistencies across sources can
lead to distorted outcomes. This challenge is even more pronounced in
datasets that span long time periods, as changes in agency structures
or reporting standards can introduce subtle biases. Long-term data
consistency is crucial for longitudinal studies and predictive
modeling, as even minor data shifts can lead to significant deviations
in model outcomes, as noted by \citet{rahm2000data} and
\citet{borgman2012conundrum}.

Finally, the intersection of data curation and machine learning for
public policy applications opens new avenues for improving governance.
High-quality data enables machine learning models to better predict
trends, allocate resources, and address service delivery issues, such
as delays in responding to 311 complaints. The COVID-19 pandemic
underscored the importance of real-time, trustworthy data in crisis
management, where the accuracy and timeliness of data-driven insights
were critical for public health responses. Without proper data
curation, machine learning models used during the pandemic would have
been compromised, affecting decisions on resource allocation and
service provision \citep{wu2020subnational, kumar2020machine}.
Future efforts could focus on automating data curation processes,
particularly in real-time data pipelines, to ensure that the data used
in machine learning models remains accurate, clean, and reliable
\citep{chu2016data, zhang2019automating}.
